On the error performance of the $A_n$ lattices

### 2012-03-08 paper.tex
more additions of reviewer comments.

\begin{keywords}
Lattices, lattice decoding, root lattice, probability of error, Voronoi cell.
\end{keywords}

@@

This paper is organised as follows.  In Section~\ref{sec:latt-latt-codes} we give a brief overview of lattices and codes constructed from them, i.e., \emph{lattice codes}.  We describe lattice decoding and show how the probability of coding error can be expressed in terms of the moments of the \emph{Voronoi cell} of the lattice used.  Section~\ref{sec:main-result} states the main result, describing recursive formula to compute the moments of the Voronoi cell of $A_n$.  Section~\ref{sec:lattice-a_n} describes the lattice $A_n$ and some of its properties.  An important property for our purposes is that the Voronoi cell of $A_n$ is precisely the orthogonal projection of a $(n+1)$-dimensional hypercube onto a hyperplace orthogonal to one of its vertices~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.
 %This result was previously known~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.  The proof is short and we give it again here so that this paper is self contained.
In Section~\ref{sec:integr-funct-over} we use this property to show how integrals over the Voronoi cell of $A_n$ can be expressed as integrals over the $(n+1)$-dimensional hypercube. These integrals are solvable and we use them to obtain the moments of the Voronoi cell in Section~\ref{sec:powers-eucl-norm}.  In Section~\ref{sec:results-simulations} we plot the probability of error versus signal to noise ratio for codes constructed from the lattices $A_1 \simeq \ints$, $A_2$, $A_3 \simeq D_3$ and $A_4$, $A_5$ and $A_8$.  We also plot the results of Monte-Carlo simulations that support our analytical results.
%An advantage of the analytical results is that the probability of error can calculated at high SNR, when the probabilty of error is small.  This is in contrast to Monte-Carlo simulations that are computationally infeasible at when the probability of error is small.  We display the performance of these lattice codes

@@

A \term{lattice}\index{lattice}, $\Lambda$, is a discrete subset of points in $\reals^m$ such that
\[
   \Lambda = \{\xbf = \Bbf\ubf \mid \ubf \in \ints^n \}
 \]
 where $\Bbf \in \reals^{m \times n}$ is an $m \times n$ matrix of rank $n$, called a \term{generator matrix} or \term{basis matrix} or simply generator or basis. In particular, the set of $n$-tuples of integers $\ints^n$ is a lattice (with the identity matrix as a generator) and we call this the~\emph{integer lattice}. A lattice $\Lambda$ associated with a rank-$n$ generator matrix $\mathbf{B}$ is said to be $n$-dimensional. If the generator is square, i.e. $m = n$, then the lattice points span $\reals^n$ and we say that the lattice is \term{full rank}. If $\Bbf$ has more rows than columns, i.e. $m > n$, then the lattice points lie in a $n$-dimensional subspace of $\reals^m$. For any lattice $\Lambda$ with an $m \times n$ generator matrix, we define $\mathcal{S}_{\Lambda}$ to be the hyperplane spanned by the columns of the generator matrix. %It is easy to see that $\mathcal{S}_\Lambda$ is then the $n$-dimensional subspace containing the lattice $\Lambda$.

The (open) \term{Voronoi cell}, denoted $\vor(\Lambda)$, of a lattice $\Lambda$ is the subset of $\mathcal{S}_{\Lambda}$ containing all points nearer (in Euclidean distance) to the lattice point at the origin than any other lattice point. The Voronoi cell is an $n$-dimensional convex polytope that is symmetric about the origin.  It is convenient to modify this definition of the Voronoi cell slightly so that the union of translated Voronoi cells $\cup_{\xbf \in \Lambda}\vor(\Lambda) + \xbf$ is equal to  $\mathcal{S}_{\Lambda}$.  That is, the Voronoi cell tessellates when translated by points in $\Lambda$.  To ensure this we require that if a face of $\vor(\Lambda)$ is open, then its opposing face is closed. Specifically, if $\xbf \in \vor(\Lambda)$ is on the boundary of $\vor(\Lambda)$ then $-\xbf \notin \vor(\Lambda)$.  We wont specifically define which opposing face is open and which is closed as the results that follow hold for any choice of open and closed opposing faces.

@@

\begin{theorem}  \label{thm:VorAn=QVorZn1}
The projection of $\vor(\ints^{n+1})$ into the zero-sum plane is equal to $\vor(A_{n})$. That is,
\[
\vor(A_n) = \Qbf\vor(\ints^{n+1}).
\]
\end{theorem}
\begin{proof}
Let $\ebf_i$ denote a vector with $i$th element equal to one and the remaining elements zero.  The $n$-volume of $\vor(A_n)$ is $\sqrt{n + 1}$.  From Burger~et.~al.~\cite[Theorem 1.1]{Burger1996} we find that the $n$-volume of the projected hypercube $\Qbf\vor(\ints^{n+1})$ is equal to
\[
\sum_{i = 1}^{n+1} \frac{\dotprod{\onebf}{\ebf_i}}{\|\onebf\|} =  \sum_{i = 1}^{n+1} \frac{1}{\sqrt{n+1}} = \sqrt{n + 1}
\]
also. It follows from Lemma~\ref{lem:QVorZnsubsetVorAn} that $\Qbf\vor(\ints^{n+1}) \subseteq \vor(A_n)$, so, because the volumes are the same, and because $\vor(A_n)$ and $\Qbf\vor(\ints^{n+1})$ are polytopes, we have $\vor(A_n) = \Qbf\vor(\ints^{n+1})$.
\end{proof}

@@

\begin{example}\emph{\textbf{(The volume of the Voronoi cell)}
In order to demonstrate this approach we will derive the $0$th moment (i.e. the volume) of the Voronoi cell using~(\ref{eq:intwewilluse}).  Setting $f(\Qbf\wbf) = \|\Qbf\wbf\|^0 = 1$ we obtain,
\begin{align*}
M_n(0) &= n\sqrt{n+1} \int^{1}_{0} \int^{y}_{0} \int^{t}_{0} \cdots \int^{t}_{0} \frac{1}{1 - t} \,dw_{n-1}\dots dw_1 \, dt  \, dy\\
 &= n\sqrt{n+1} \int^{y}_{0} \int_{0}^{1}  \frac{t^{n-1}}{1 - t} \, dt  \, dy \\
&= n\sqrt{n+1}  \int^{1}_{0} \beta(y, n, 0) \, dy = \sqrt{n+1},
\end{align*}
as required.  Here $\beta(x,a,b) =  \int_{0}^{x}  t^{a-1}(1 - t)^{b-1} \,dt$ is the incomplete beta function~\cite{Pearson_tables_of_beta_functions} and we have used the identity $\int^{1}_{0} \beta(y, n, 0) \, dy = \frac{1}{n}$.}
\end{example}

>>>>

\begin{IEEEkeywords}
Lattices, lattice decoding, root lattice, probability of error, Voronoi cell.
\end{IEEEkeywords}

@@

This paper is organised as follows.  In Section~\ref{sec:latt-latt-codes} we give a brief overview of lattices and codes constructed from them, i.e., \emph{lattice codes}.  We describe lattice decoding and show how the probability of coding error can be expressed in terms of the moments of the \emph{Voronoi cell} of the lattice used.  Section~\ref{sec:main-result} states the main result, describing recursive formula to compute the moments of the Voronoi cell of $A_n$.  Section~\ref{sec:lattice-a_n} describes the lattice $A_n$ and some of its properties.  An important property for our purposes is that the Voronoi cell of $A_n$ is precisely the orthogonal projection of an $(n+1)$-dimensional hypercube onto a hyperplace orthogonal to one of its vertices~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.
 %This result was previously known~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.  The proof is short and we give it again here so that this paper is self contained.
In Section~\ref{sec:integr-funct-over} we use this property to show how integrals over the Voronoi cell of $A_n$ can be expressed as integrals over the $(n+1)$-dimensional hypercube. These integrals are solvable and we use them to obtain the moments of the Voronoi cell in Section~\ref{sec:powers-eucl-norm}.  In Section~\ref{sec:results-simulations} we plot the probability of error versus signal to noise ratio for codes constructed from the lattices $A_1 \simeq \ints$, $A_2$, $A_3 \simeq D_3$ and $A_4$, $A_5$ and $A_8$.  We also plot the results of Monte-Carlo simulations that support our analytical results.
%An advantage of the analytical results is that the probability of error can calculated at high SNR, when the probabilty of error is small.  This is in contrast to Monte-Carlo simulations that are computationally infeasible at when the probability of error is small.  We display the performance of these lattice codes

@@

A \term{lattice}\index{lattice}, $\Lambda$, is a discrete subset of points in $\reals^m$ such that
\[
   \Lambda = \{\xbf = \Bbf\ubf \mid \ubf \in \ints^n \}
 \]
 where $\Bbf \in \reals^{m \times n}$ is an $m \times n$ matrix of rank $n$, called a \term{generator matrix} or \term{basis matrix} or simply generator or basis. In particular, the set of $n$-tuples of integers $\ints^n$ is a lattice (with the identity matrix as a generator) and we call this the~\emph{integer lattice}. A lattice $\Lambda$ associated with a rank-$n$ generator matrix $\mathbf{B}$ is said to be $n$-dimensional. If the generator is square, i.e. $m = n$, then the lattice points span $\reals^n$ and we say that the lattice is \term{full rank}. If $\Bbf$ has more rows than columns, i.e. $m > n$, then the lattice points lie in an $n$-dimensional subspace of $\reals^m$. For any lattice $\Lambda$ with an $m \times n$ generator matrix, we define $\mathcal{S}_{\Lambda}$ to be the hyperplane spanned by the columns of the generator matrix. %It is easy to see that $\mathcal{S}_\Lambda$ is then the $n$-dimensional subspace containing the lattice $\Lambda$.

The (open) \term{Voronoi cell}, denoted $\vor(\Lambda)$, of a lattice $\Lambda$ is the subset of $\mathcal{S}_{\Lambda}$ containing all points nearer (in Euclidean distance) to the lattice point at the origin than any other lattice point. The Voronoi cell is an $n$-dimensional convex polytope that is symmetric about the origin.  It is convenient to modify this definition of the Voronoi cell slightly so that the union of translated Voronoi cells $\cup_{\xbf \in \Lambda}\vor(\Lambda) + \xbf$ is equal to  $\mathcal{S}_{\Lambda}$.  That is, the Voronoi cell tessellates when translated by points in $\Lambda$.  To ensure this we require that if a face of $\vor(\Lambda)$ is open, then its opposing face is closed. Specifically, if $\xbf \in \vor(\Lambda)$ is on the boundary of $\vor(\Lambda)$ then $-\xbf \notin \vor(\Lambda)$.  We will not specifically define which opposing face is open and which is closed as the results that follow hold for any choice of open and closed opposing faces.

@@

\begin{theorem}  \label{thm:VorAn=QVorZn1}
The projection of $\vor(\ints^{n+1})$ into the zero-sum plane is equal to $\vor(A_{n})$. That is,
\[
\vor(A_n) = \Qbf\vor(\ints^{n+1}).
\]
\end{theorem}
\begin{IEEEproof}
Let $\ebf_i$ denote a vector with $i$th element equal to one and the remaining elements zero.  The $n$-volume of $\vor(A_n)$ is $\sqrt{n + 1}$.  From Burger~et.~al.~\cite[Theorem 1.1]{Burger1996} we find that the $n$-volume of the projected hypercube $\Qbf\vor(\ints^{n+1})$ is equal to
\[
\sum_{i = 1}^{n+1} \frac{\dotprod{\onebf}{\ebf_i}}{\|\onebf\|} =  \sum_{i = 1}^{n+1} \frac{1}{\sqrt{n+1}} = \sqrt{n + 1}
\]
also. It follows from Lemma~\ref{lem:QVorZnsubsetVorAn} that $\Qbf\vor(\ints^{n+1}) \subseteq \vor(A_n)$, so, because the volumes are the same, and because $\vor(A_n)$ and $\Qbf\vor(\ints^{n+1})$ are polytopes, we have $\vor(A_n) = \Qbf\vor(\ints^{n+1})$.
\end{IEEEproof}

@@

\begin{example}\emph{\textbf{(The volume of the Voronoi cell)}
In order to demonstrate this approach we will derive the $0$th moment (i.e. the volume) of the Voronoi cell using~(\ref{eq:intwewilluse}).  Setting $f(\Qbf\wbf) = \|\Qbf\wbf\|^0 = 1$ we obtain,
\begin{align*}
M_n(0) &= n\sqrt{n+1} \int^{1}_{0} \int^{y}_{0} \int^{t}_{0} \cdots \int^{t}_{0} \frac{1}{1 - t} \,dw_{n-1}\dots dw_1 \, dt  \, dy\\
 &= n\sqrt{n+1} \int_{0}^{1} \int^{y}_{0} \frac{t^{n-1}}{1 - t} \, dt  \, dy \\
&= n\sqrt{n+1}  \int^{1}_{0} \beta(y, n, 0) \, dy = \sqrt{n+1},
\end{align*}
as required.  Here $\beta(x,a,b) =  \int_{0}^{x}  t^{a-1}(1 - t)^{b-1} \,dt$ is the incomplete beta function~\cite{Pearson_tables_of_beta_functions} and we have used the identity $\int^{1}_{0} \beta(y, n, 0) \, dy = \frac{1}{n}$.}
\end{example}
